### 2017-02-28 intro.tex
Added James' specific/easy comments

\section{The Cosmological Context of Galaxy Formation}
The modern picture of galaxy formation sits within the context of a $\Lambda$
Cold Dark Matter  ($\Lambda CDM$) cosmology
\citep{Rees1977,White1978,Blumenthal1984,Mo1998}.  This is a universe which has
a no curvature ($\Omega = 1$), and today has the majority of its mass-energy in
a cosmological constant $\Lambda$ ($\Omega_\Lambda \sim 0.7$), and most of its
matter in a collisionless form (dark matter) ($\Omega_m \sim 0.3$)
\citep{Planck2014}.  Roughly $5\%$ of the matter-energy density of the universe
exists as baryons ($\Omega_b \sim 0.05$), which existed as a nearly isothermal
($\delta T/T\sim10^{-5}$) gas of hydrogen and helium at the earliest
observational time.  The composition of the universe, and the precise value of
the mass-energy density for these components sets the lifetime of the universe,
how rapidly it expands (or contracts), and has tremendous impact on the growth
of structure.  This structure is ultimately where galaxies will form. All of
this began $13.8$ billion years ago, with the universe emerging from a state of
high temperature and density, expanding to form what we see today.  The
primordial elemental abundances, with $\sim4/5$ of the universe's baryons in
hydrogen, and the remainder in helium, was established just a few minutes after
the big bang, during a brief period of ``Big Bang Nucleosynthesis''
\citep{Alpher1948}.  The remaining heavier elements have been formed via stellar
nucleosynthesis, through fusion in the cores of stars and in energetic processes
that take place during supernovae (SN) \citep{Wagoner1967}, or through the
capture of high energy cosmic ray protons and $\alpha$ particles by carbon, nitrogen, or
oxygen nucleii \citep{Reeves1970}.  The chemical composition of the universe
establishes the kinds of stars that form within galaxies and the rates at which
gas within the ISM and beyond cools.

Much of what we know about the cosmology of our universe comes from observations
of the first light to escape the universe's surface of last scattering as it
became optically thin: the cosmic microwave background radiation (CMB).  Prior
to a redshift of $z\approx1100$, the universe had a high ionization fraction
such that the mean free path for photons was short, making all of space
effectively opaque.  Once the universe became cool enough that hydrogen became
fully neutral, the optical depth fell low enough that photons could free-stream,
decoupled from baryons.  This ultraviolet light we observe today, redshifted to
microwaves with a temperature of $\sim2.7\;\rm{K}$.  When the power spectrum of
this radiation is measured \citep{Spergel2003}, features in that spectrum can be
used to determine the components of the universe (with some ambiguity that can
be resolved via SN Ia and large scale structure measurements
\citealt{Riess1998,Perlmutter1999,Beutler2011,Blake2011}), as well as the
spectrum of density perturbations that will eventually form galaxies
\citep{Press1974,Peebles1980}.   The CMB is the single most important
observational tool for understanding the cosmology of our universe.

The formation of galaxies within the universe began with the gravitational
collapse of small, linear density perturbations that were likely seeded by
quantum fluctuations amplified by inflation \citep{Guth1981,Linde1982}.  These
perturbations grow through gravitational collapse.  Until $z\sim1100$, only dark
matter perturbations were able to grow.  Prior to this, baryons were coupled to
photons, causing density perturbations to oscillate as stable sound waves.  This
early collapse of matter meant that when decoupling occurred at $z\sim1100$,
baryons were able to collapse into the deeper potential wells grown out of
dark-matter density perturbations.  These overdensities expand slower that the
surrounding medium, as they experience a drag from their own self-gravity.
Eventually, these regions stop expanding altogether and begin collapsing at the
turnaround time $t_{turn}$, and virialize at $2\;t_{turn}$ \citep{Peebles1980}.
The first galaxies likely began to form around $z\sim20-50$.  Without the
earlier growth of dark matter perturbations, structure formation would have
occurred much later, and the number of galaxies and distribution of their masses
would be different than those we see today \citep{Davis1985}.

The perturbation spectrum that forms the structure in which galaxies begins to
form is essentially a Gaussian random field, which \citet{Press1974} were able
to show produces a halo mass spectrum at a given time with a form that matched
the functional form for the observed galaxy mass function
\citet{Schechter1976}:
\begin{equation}
    \frac{dn}{dM}(M,t) =
    \sqrt{\frac{2}{\pi}}\frac{\rho_0\delta_c(t)}{M^2\sigma(M)} 
    \left\lvert\frac{d\ln{\sigma}}{d\ln{M}}\right\rvert
    \exp{\left[-\frac{\delta^2_c(t)}{2\sigma^2(M)}\right]}
\end{equation}
In this equation, $\rho_0$ is the mean density of the universe, and
$\delta_c(t)$ is the critical overdensity for collapse at a given time $t$.
The mass variance spectrum, $\sigma^2(M)$ can be determined observationally
using the CMB power spectrum \citep{Fixsen1996,Spergel2003,Planck2015} and an
assumption of a specific cosmology.  This function gives a distribution of
masses for the dark matter halos that galaxies will begin to form within.  As
these halos collapse, gravitational tugs from their surroundings will impart
torques on them \citep{Barnes1987}, causing halos to have a distribution of
angular momenta.  High resolution N-body simulations have found that dark
matter halos have both a universal density profile
\citep{Navarro1996,Merritt2006} as well as a universal angular momentum profile
\citep{Bullock2001}.  These dark matter halos all serve as the environment in
which the collapse of baryons and formation of stars take place.

The collapse of baryons within halos will depend strongly on whether the
post-virialized gas can cool radiatively \citep{Rees1977}.  As gas accretes onto
a halo with overdensity $\Delta$ it passes through a shock that raises its
temperature to the virial temperature:
\begin{equation}
    T_{virial} = \frac{2}{3}\frac{G\mu m_H}{k_B}\left({\frac{4\pi\rho_0\Delta
    M^2}{3}}\right)^{1/3}
\end{equation}
This means that more massive halos have larger virial temperatures, and,
depending on the mode of radiative cooling (atomic cooling from primordial
hydrogen, molecular hydrogen cooling, or cooling through metal lines), this may
give cooling times shorter than the current age of the universe.  These halos
are the ones in which galaxy formation may begin, as their hot gaseous halos can
collapse to form a central concentration where star formation can occur.  In
some halos, these virial shocks may be unstable to cooling, allowing the direct
collapse of cold gas onto the disc \citep{Birnboim2003}.

As gas does begin to cool within these halos, it will lose pressure support (by
radiating away thermal energy), but {\it not} angular momentum.  For large
spiral galaxies, this means that the baryons, which typically carry the same
specific angular momentum as the halo itself, will flatten to form an
exponential disc as they cool \citep{Fall1980,Mo1998}.  Streams of gas
accreting along the cosmic web may also provide angular momentum to the forming
disc \citep{Dekel2006}.  Within this disc, an ISM will form.  This ISM is
multiphase, containing both hot ($T\sim10^4\;\rm{K}$) and cold
($T\sim10-100\;\rm{K}$) gas in quasistatic equilibrium \citep{McKee1977}.  This
multiphase state is highly dependent on both the ISM metallicity and the local
heating rates (which themselves can vary with redshift)
\citep{Field1969,Wolfire1995,Norman1997}.
\begin{figure}
    \includegraphics[width=0.7\textwidth]{mass_function.eps}
    \caption[Galaxy mass function]{The observed mass function for galaxies in
    the nearby universe (solid line) does not match the mass function for dark
    matter halos given in equation 1, scaled by the baryon fraction (dashed
    line).  This tells us that there are additional physical processes involved
    in the formation of galaxies within their dark matter halos.  The most
    likely candidate for this is feedback from stars and black holes.  Of
    particular interest is the peak in star formation efficiency near
    $M_{200}=10^{12}M_\odot$.  \textit{Image credit: Adapted from figure 1 in
    \citet{Ferrero2012}}}.
\end{figure}

@@

In the intervening 75 years, we have migrated to much less labor-intensive
simulations, relying on the constantly growing power of digital computers.  As
Gordon Moore first observed in 1965, the processing power of computers has been
continuously growing at an exponential rate for more than half a century.  This
has meant that, even without advances in numerical methods and algorithm design,
the ability of simulators today to model a system in high resolution is
monumentally greater than just 10 years ago.

@@

The earliest numerical simulations of galaxy formation focused on the effects of
gravity alone.  A classic example of this are the simulations presented in
\citet{Toomre1972}, which showed that the irregular structures seen in galaxies
like NGC 4676 (``The Mice'') or NGC4038/4039 (``The Antennae Galaxies'') are the
results of major mergers between two disc galaxies, as is clear from figure 1.2.
These simulations used discs composed of a mere 120 test particles, and ignored
the effects of self-gravity between these particles.  Despite their crudeness,
these simulations definitively showed that the morphologies of galactic bridges and tails could be
formed by the tidal interactions of two passing/merging galaxies.  The IBM
360/95 that \citet{Toomre1972} (with an inflation-adjusted purchase price in the
millions of dollars) used for their simulations was capable of 5.5 million
instructions per second, roughly $0.2\%$ the speed of a 35\$ Raspberry Pi 3
available today.

Modern galaxy simulations do far more than merely integrate the equations of
motion for a few hundred test particles in a gravitational field.  Today, state
of the art simulations will track gravity (including self-gravity) for stars,
gas, and dark matter \citep{Aarseth1980,Stadel2001,Dehnen2002}; evolve Euler's
equations for the hydrodynamics of gaseous baryons
\citep{Wadsley2004,Teyssier2002,Bryan2014}; include models for radiative
cooling of gas that allow for cooling from both hydrogen and metal lines
\citep{Marri2003,Shen2010}; form stars from dense gas
\citep{Katz1992,Agertz2015}; and include models for feedback from massive stars
and black holes \citep{Katz1992,Springel2003,DiMatteo2005}.  Often, simulations
will also include treatments for radiative transfer
\citep{Gendelev2012,Krumholz2013} and magneto-hydrodynamics
\citep{Girichidis2015} as well.

@@

Gravity is typically solved using algorithms designed to minimize the workload
that would be needed for a direct summation of each particle or grid cell
($\mathcal{O}(N^2)$ for N grid cells/particles).  This becomes intractable for
larger numbers of resolution elements, as the workload will scale as the sixth
power of the linear resolution (an order of magnitude better linear resolution
will be $10^6$ times more expensive!).  In order to reduce these costs,
algorithms have been developed that allow for more efficient gravitational
force calculations.  Tree-based methods \citep{Barnes1986} reduce the
computational cost of long-ranged (and thus less significant) force
calculations by using space-partitioning trees.  Short-range force calculations
are done particle-by-particle, while long-range forces are calculated against
larger tree cells, thus reducing the cost of calculating gravity to
$\mathcal{O}(N\log N)$.  Particle-Particle/Particle-Mesh ($P^3M$) methods
\citep{Couchman1991} produce similar scalings, but rather than using a
spatial tree rely on a spectral Fourier-based approach to calculate long-ranged
interactions.  Recently, the Fast Multipole Method \citep{Greengard1987} has
been applied to the calculation of gravitational forces
\citep{Dehnen2002,Hahn2013} using a multipole expansion of Green's functions.
This allows for extremely efficient gravity calculation, scaling linearly with
particle number ($\mathcal{O}(N)$).

Hydrodynamical solvers usually come in one of two classes: Eulerian or
Lagrangian.  Eulerian methods discretize space into fixed volume elements with
static boundaries.  By breaking a simulation volume into a grid of cells that
each contains scalar and vector quantities those quantities can be advected
between cells according to the solver scheme.  Eulerian schemes typically allow
for higher-order solvers, lower intrinsic noise, and better resolved
discontinuities \citep{Teyssier2002,Stone2008,Bryan2014}.  They suffer from
intrinsic mixing and diffusion that scales with the flow velocity relative to
the grid, poor integration of circular orbits, and costly timestep limits
\citep{Agertz2007,Tasker2008}.  Lagrangian methods instead discretize mass,
breaking the gaseous volume into discrete points, each of which moves at the
fluid velocity to advect the quantities it contains, with smoothing functions
used to interpolate values in between particles; hence the name for the most
common Lagrangian method, Smoothed Particle Hydrodynamics (SPH).  Lagrangian
methods couple easily to tree codes for gravity calculation, are Galilean
invariant, and intrinsically conserve mass, energy, and momentum (linear and
angular) \citep{Katz1996,Wadsley2004,Springel2005}.  SPH requires the use of
artificial viscosity to handle shocks, and as such have difficulty in capturing
subsonic turbulence \citep{Bauer2012}. Recently, new methods have begin to
hybridize features of each of these classes as well
\citep{Springel2010,Hopkins2015}.

@@

Beyond the choice of algorithms, the choice of what type of simulation to set up
in the initial conditions is also important.  To study large populations of
galaxies, along with large-scale structure, large-volume cosmological boxes with
volumes $\gtrsim 10^6\;rm{Mpc^3}$, but with spatial resolutions $\sim \rm{kpc}$
are used.  Studying how individual galaxies form is typically done with
cosmological zoom simulations, where a comparatively cheap dark-matter only
simulation is first run to select a region of interest, which is then
resimulated with higher resolution and baryonic physics included
\citep{Navarro1993}. Zoom-in simulations can typically achieve an order of
magnitude or better resolution than full cosmological boxes.  Isolated galaxy
disc simulations cannot show how galaxies form in a cosmological environment,
but can be run with $\sim \rm{pc}$ resolution to study closely how internal
processes within a galaxy take place \citep{Hopkins2011,Benincasa2016}.  These
details of these internal processes are discussed in section 1.3.

The earliest attempt to track the formation of a galaxy in a cosmological
context with the effects of star formation and feedback included was by
\citet{Katz1992}.  This simulation used the Lagrangian TREESPH method
\citep{Hernquist1989} to evolve a parcel of gas and dark matter from an initial
linear perturbation, including small-scale power derived from CDM predictions
\citep{Zeldovich1970,Peebles1982}.  Gas was allowed to cool radiatively using
rates calculated for primordial gas.  This study introduced a model for star
formation that is still the standard for most simulations of galaxy formation.
Observations \citep{Kennicutt1998} have suggested that the star formation rate
is a power law of the gas density: $\dot\Sigma_*\propto\Sigma_{gas}^\alpha$, with
$\alpha=1-2$.  If it is assumed that gas collapses to form stars in some
multiple $c_*$ of its freefall time \citep{Schmidt1959}, $t_{ff}$, a simple
3-dimensional star formation model can be built with the form:
\begin{equation}
    \dot\rho_* = \frac{c_*\rho_{gas}}{t_{ff}} = c_*\sqrt{\frac{\rho^3_{gas}}{4\pi G}}
\end{equation}
\citet{Katz1992} used a $c_*$ parameter of $0.1$, which is within the range of
what is typically used in simulations today ($1-100\%$).  On top of the model
for star formation, feedback was included without the use of any significant
subgrid model:  $10^{51}\;\rm{erg/SN}$ worth of energy was injected back into
the ISM by smoothing it over a number of resolution elements.  One of the key
findings of \citet{Katz1992} was that the vast majority of SN energy was
radiated away, making them very ineffective at regulating star formation within
the galaxy.  In the end, \citet{Katz1992} found that the peak star formation in
the simulated galaxy was nearly 10 times greater than in observed galaxies.
\citet{Thacker2000} later showed that the smoothing of energy used by
\citet{Katz1992} resulted in a large overestimation of the cooling rates in
feedback-heated gas. Since then, much work has gone into developing models for
SN feedback that address the issue of overcooling that was present in these
earlier simulations.

@@

\subsection{Superbubbles: The Natural Scale for Feedback}
The formation of massive stars takes place mostly in clusters of $10^4$ or more
stars.  This means that the SN of these stars can be treated not as
individual dumps of $10^{51}\;\rm{erg/SN}$, but instead as a luminosity of
$10^{34}\;\rm{\frac{erg}{s \cdot M_\odot}}$.  This allows us to use stellar wind
models like \citet{Weaver1977} to study the evolution of superbubbles driven by
clustered SN.  Each individual SN's ejecta thermalizes in a small
region within the center of the bubble, and the resulting hot gas drives a shock
that sweeps up the surrounding ISM.  In a short time, the swept-up ISM can
radiatively cool efficiently , resulting in a thin, cold shell surrounding a
hot, diffuse bubble.

>>>>

\section{The Cosmological Context of Galaxy Formation} The modern picture of
galaxy formation sits within the context of a $\Lambda$ Cold Dark Matter
($\Lambda CDM$) cosmology \citep{Rees1977,White1978,Blumenthal1984,Mo1998}.
This is a universe which has a no curvature ($\Omega = 1$), and today has the
majority of its mass-energy in dark energy $\Lambda$ ($\Omega_\Lambda =
0.6911\pm0.0062$), and most of its matter in a collisionless form (dark matter)
($\Omega_m = 0.3089\pm0.0062$) \citep{Planck2015b}.  Roughly $5\%$ of the
matter-energy density of the universe exists as baryons ($\Omega_b =
0.048613\pm0.00031$).  The composition of the universe, and the precise value of
the mass-energy density for these components sets the lifetime of the universe,
how rapidly it expands (or contracts), and exactly determines the growth of
large-scale structure.  Within this structure is ultimately where galaxies will
form.  All of this began $13.799\pm0.021$ billion years ago \citep{Planck2015b},
with the universe emerging from a state of high temperature and density,
expanding to form what we see today.  The primordial elemental abundances, with
$\sim4/5$ of the universe's baryons in hydrogen, and the remainder in helium,
were established just a few minutes after the big bang, during a brief period of
``Big Bang Nucleosynthesis'' \citep{Alpher1948}.  The remaining heavier elements
have been formed via stellar nucleosynthesis, through fusion in the cores of
stars and in energetic processes that take place during supernovae (SN)
\citep{Wagoner1967}, or through the capture of high energy cosmic ray protons
and $\alpha$ particles by carbon, nitrogen, or oxygen nucleii
\citep{Reeves1970}.  The chemical composition of the universe establishes the
kinds of stars that form within galaxies and the rates at which gas within the
interstellar medium (ISM) and beyond cools.

Much of what we know about the cosmology of our universe comes from observations
of the first light to escape the universe's surface of last scattering as it
became optically thin: the cosmic microwave background radiation (CMB).  
When the CMB was emitted, the baryons of the universe were in the form of a
nearly isothermal ($\delta T/T\sim10^{-5}$) gas of hydrogen and helium.  Prior
to a redshift of $z\approx1100$, the universe had a high ionization fraction
such that the mean free path for photons was short, making all of space
effectively opaque.  Once the universe became cool enough that hydrogen became
fully neutral, the optical depth fell low enough that photons could free-stream,
decoupled from baryons.  We observe this ultraviolet light today, redshifted to
microwaves with a temperature of $2.729\pm0.004\;\rm{K}$ \citep{Fixsen1996}.
When the power spectrum of this radiation is measured \citep{Spergel2003},
features in that spectrum can be used to determine the components of the
universe (with some degeneracy that can be resolved via SN Ia and large scale
structure measurements
\citealt{Riess1998,Perlmutter1999,Beutler2011,Blake2011}). These features also
reveal the spectrum of density perturbations that will eventually form galaxies
\citep{Press1974,Peebles1980}.   The CMB is the single most important
observational tool for understanding the cosmology of our universe.

The formation of structure within the universe began with the gravitational
collapse of small, linear density perturbations seeded by quantum fluctuations
amplified by inflation \citep{Guth1981,Linde1982}.  These perturbations grow
through gravitational collapse.  Until $z\sim1100$, only dark matter
perturbations were able to grow.  Prior to this, baryons were coupled to
photons, causing density perturbations to oscillate as stable sound waves.  This
early collapse of matter meant that when decoupling occurred at $z\sim1100$,
baryons were able to collapse into the deeper potential wells grown out of
dark-matter density perturbations.  These overdensities expand slower that the
surrounding medium, as they experience a drag from their own self-gravity.
Eventually, these regions stop expanding and begin collapsing at the turnaround
time $t_{turn}$, and virialize at roughly $2\;t_{turn}$ \citep{Peebles1980}.
The first galaxies likely began to form around $z\sim20-50$, with minimum sizes
set by the baryonic physics of Silk damping and radiative cooling
\citep{Silk1968}.  Without the earlier growth of dark matter perturbations,
structure formation would have occurred much later, and the number of galaxies
and distribution of their masses would be different than those we see today
\citep{Davis1985}.

The density perturbation spectrum describes the complete statistics of the
Gaussian random field from which structure forms. \citet{Press1974} were able to
show this produces a halo mass spectrum at a given time with a form that matched
the functional form for the observed galaxy mass function
\citet{Schechter1976}:
\begin{equation}
    \frac{dn}{dM}(M,t) =
    \sqrt{\frac{2}{\pi}}\frac{\rho_0\delta_c(t)}{M^2\sigma(M)} 
    \left\lvert\frac{d\ln{\sigma}}{d\ln{M}}\right\rvert
    \exp{\left[-\frac{\delta^2_c(t)}{2\sigma^2(M)}\right]}
\end{equation}
In this equation, $\rho_0$ is the mean density of the universe, and
$\delta_c(t)$ is the critical overdensity for collapse at a given time $t$.  The
mass variance spectrum, $\sigma^2(M)$ can be determined observationally using
the CMB power spectrum \citep{Fixsen1996,Spergel2003,Planck2015} and an
assumption of a specific cosmology.  This function gives an approximate
distribution of masses for the dark matter halos that galaxies will begin to
form within.  As these halos collapse, non-linear gravitational tugs from their
surroundings will impart torques on them \citep{Barnes1987}, causing halos to
have a distribution of angular momenta.  High resolution N-body simulations have
found that dark matter halos have both a universal density profile
\citep{Navarro1996,Merritt2006} as well as a universal angular momentum profile
\citep{Bullock2001}.  These dark matter halos serve as the environment in
which the collapse of baryons and formation of stars take place, fed by
accretion along large-scale filaments.

The collapse of baryons within halos will depend strongly on whether the
post-virialized gas can cool radiatively \citep{Rees1977}.  As gas accretes onto
a halo with overdensity $\Delta$ it collisionally converts its potential energy
into heat, with a characteristic virial temperature:
\begin{equation}
    T_{virial} = \frac{2}{3}\frac{G\mu m_H}{k_B}\left({\frac{4\pi\rho_0\Delta
    M^2}{3}}\right)^{1/3}
\end{equation}
This means that more massive halos have larger virial temperatures, and,
depending on the mode of radiative cooling (atomic cooling from primordial
hydrogen, molecular hydrogen cooling, or cooling through metal lines), this may
give cooling times shorter than the current age of the universe.  These halos
are the ones in which galaxy formation may begin, as their hot gaseous halos can
collapse to form a central concentration where star formation can occur.  In
some halos, virial shocks may be unstable to cooling, allowing the direct
collapse of cold gas onto the disc \citep{Birnboim2003}.

As gas does begin to cool within these halos, it will lose pressure support (by
radiating away thermal energy), but {\it not} angular momentum.  For large
spiral galaxies, this means that the baryons, which typically carry the same
specific angular momentum as the halo itself, will flatten to form an
exponential disc as they cool \citep{Fall1980,Mo1998}.    A combination
of heating and cooling processes will render this gas multiphase, containing
both warm/hot ($T>10^4\;\rm{K}$) and cold ($T\sim10-100\;\rm{K}$) gas in
quasistatic equilibrium \citep{McKee1977}.  This multiphase state is highly
dependent on both the metallicity and the local heating rates (which
themselves can vary with redshift) \citep{Field1969,Wolfire1995,Norman1997}.  In
the disc of the galaxy, this dense multiphase gas forms the ISM from which stars
collapse and are born.
\begin{figure}
    \includegraphics[width=0.7\textwidth]{mass_function.eps}
    \caption[Galaxy mass function]{The observed mass function for galaxies in
    the nearby universe (derived using abundance matching for the solid line and
    semi-analytic modelling for the green points) does not match the mass
    function for dark matter halos given in equation 1, scaled by the baryon
    fraction (dashed line).  This tells us that there are additional physical
    processes involved in the formation of galaxies within their dark matter
    halos.  The most likely candidate for this is feedback from stars and black
    holes.  Of particular interest is the peak in star formation efficiency near
    $M_{200}=10^{12}M_\odot$.  \textit{Image credit: Adapted from figure 1 in
    \citet{Ferrero2012}}}.
\end{figure}

@@

In the intervening 75 years, we have migrated to much less labor-intensive
simulations, relying on the constantly growing power of digital computers.  As
Gordon Moore first observed in 1965, the processing power of computers has been
continuously growing at an exponential rate for more than half a century.  This
has meant that, even without advances in numerical methods and algorithm design,
the ability of simulators today to model a system in high resolution is
monumentally greater than just 10 years ago.  Beyond the simple improvement of
hardware, significant advances have been made in the underlying algorithms used
in simulations, allowing the same hardware to perform larger, higher resolution
simulations than ever before.  These algorithmic improvements have increased the
speed and scalability of simulation codes even faster than improvements in
hardware.

@@

The earliest numerical simulations of galaxy formation focused on the effects of
gravity alone.  A classic example of this are the simulations presented in
\citet{Toomre1972}, which showed that the irregular structures seen in galaxies
like NGC 4676 (``The Mice'') or NGC4038/4039 (``The Antennae Galaxies'') are the
results of major mergers between two disc galaxies, as is clear from figure 1.2.
These simulations used discs composed of a mere 120 test particles, and ignored
the effects of self-gravity between these particles.  Despite their crudeness,
these simulations definitively showed that the morphologies of galactic bridges and tails could be
formed by the tidal interactions of two passing/merging galaxies.  The IBM
360/95 that \citet{Toomre1972} (with an inflation-adjusted purchase price in the
millions of dollars) used for their simulations was capable of 5.5 million
instructions per second, roughly $0.2\%$ the speed of a \$35 Raspberry Pi 3
available today.

Modern galaxy simulations do far more than merely integrate the equations of
motion for a few hundred test particles in a gravitational field.  Today, state
of the art simulations will track gravity (including self-gravity) for stars,
gas, and dark matter \citep{Aarseth1980,Stadel2001,Dehnen2002}; evolve Euler's
equations for the hydrodynamics of gaseous baryons
\citep{Wadsley2004,Teyssier2002,Bryan2014}; include models for radiative cooling
of gas that allow for cooling from both hydrogen and metal lines
\citep{Marri2003,Shen2010}; form stars from dense gas
\citep{Katz1992,Agertz2015}; and include models for feedback from massive stars
and black holes \citep{Katz1992,Springel2003,DiMatteo2005}.  These feedback
models have been the focus of much research in the past decade, as we shall
discuss further in this work.  Simulations also include treatments
for radiative transfer \citep{Gendelev2012,Krumholz2013} and
magneto-hydrodynamics \citep{Girichidis2015} as well.

@@

\subsection{Gravity}
Gravity is typically solved using algorithms designed to minimize the workload
that would be needed for a direct summation of each particle or grid cell
($\mathcal{O}(N^2)$ for N grid cells/particles).  This becomes intractable for
larger numbers of resolution elements, as the workload will scale as the sixth
power of the linear resolution (an order of magnitude better linear resolution
will be $10^6$ times more expensive!).  In order to reduce these costs,
algorithms have been developed that allow for more efficient gravitational
force calculations.  Tree-based methods \citep{Barnes1986} reduce the
computational cost of long-ranged (and thus less significant) force
calculations by using space-partitioning trees.  Short-range force calculations
are done particle-by-particle, while long-range forces are calculated against
larger tree cells, thus reducing the cost of calculating gravity to
$\mathcal{O}(N\log N)$.  Particle-Particle/Particle-Mesh ($P^3M$) methods
\citep{Couchman1991} produce similar scalings, but rather than using a
spatial tree rely on a spectral Fourier-based approach to calculate long-ranged
interactions.  Recently, the Fast Multipole Method \citep{Greengard1987} has
been applied to the calculation of gravitational forces
\citep{Dehnen2002,Hahn2013} using a multipole expansion of Green's functions.
This allows for extremely efficient gravity calculation, scaling linearly with
particle number ($\mathcal{O}(N)$).  This method allows it to now be possible to
run N-body simulations with greater than $10^{12}$ particles\citep{Potter2016}.

\subsection{Hydrodynamics}
Hydrodynamical solvers usually come in one of two classes: Eulerian or
Lagrangian.  Eulerian methods discretize space into fixed volume elements with
fixed boundaries.  By breaking a simulation volume into a grid of cells that
each contains scalar and vector quantities those quantities can be advected
between cells according to the solver scheme.  Eulerian schemes typically allow
for higher-order solvers, lower intrinsic noise, and better resolved
discontinuities \citep{Teyssier2002,Stone2008,Bryan2014}.  They suffer from
intrinsic mixing and diffusion that scales with the flow velocity relative to
the grid, poor integration of circular orbits, and costly timestep limits
\citep{Agertz2007,Tasker2008}.  Lagrangian methods instead discretize mass,
breaking the gaseous volume into discrete points, each of which moves at the
fluid velocity to advect the quantities it contains, with smoothing functions
used to interpolate values in between particles; hence the name for the most
common Lagrangian method, Smoothed Particle Hydrodynamics (SPH).  Lagrangian
methods couple easily to tree and particle-mesh codes for gravity calculation,
are Galilean invariant, and intrinsically conserve mass, energy, and momentum
(linear and angular) \citep{Katz1996,Wadsley2004,Springel2005}.  The perfect
conservation of angular momentum is critical for accurately tracking orbits. SPH
requires the use of artificial viscosity to handle shocks, and as such have
difficulty in capturing subsonic turbulence \citep{Bauer2012}. Recently, new
methods have begin to hybridize features of each of these classes as well
\citep{Springel2010,Hopkins2015}.

@@

\subsection{Designing Simulations}
Beyond the choice of algorithms, the construction of appropriate inital
conditions is also important.  To study large populations of galaxies, along
with large-scale structure, large-volume cosmological boxes with volumes
$\gtrsim 10^6\;\rm{Mpc^3}$, but with spatial resolutions $\sim \rm{kpc}$ are
used (for example, the {\it Illustris} simulations have a softening length of
$1.4\;rm{kpc}$ \citealt{Vogelsberger2014}).  Studying how individual galaxies
form is typically done with cosmological zoom simulations, where a comparatively
cheap dark-matter only simulation is first run to select a region of interest,
which is then resimulated with higher resolution and baryonic physics included
\citep{Navarro1993}. Zoom-in simulations can typically achieve an order of
magnitude or better resolution than full cosmological boxes.  Isolated galaxy
disc simulations use a constructed disc profiles containing gas and stars,
embedded within a dark halo. These cannot show how galaxies form in a
cosmological environment, but can be run with $\sim \rm{pc}$ resolution to study
closely how internal processes within a galaxy take place
\citep{Hopkins2011,Benincasa2016}.  Details of these internal processes are
discussed in section 1.3.

The earliest attempt to track the formation of a galaxy in a cosmological
context with the effects of star formation and feedback included was by
\citet{Katz1992}.  This simulation used the Lagrangian TREESPH 
\citep{Hernquist1989} to evolve a volume of gas and dark matter from an initial
linear perturbation, including small-scale power derived from CDM predictions
\citep{Zeldovich1970,Peebles1982}.  Gas was allowed to cool radiatively using
rates calculated for primordial gas.  This study introduced a model for star
formation that is still the standard for most simulations of galaxy formation.
Observations \citep{Kennicutt1998} have suggested that the star formation rate
is a power law of the gas density: $\dot\Sigma_*\propto\Sigma_{gas}^\alpha$, with
$\alpha=1-2$.  If it is assumed that gas collapses to form stars in some
multiple $c_*$ of its freefall time \citep{Schmidt1959}, $t_{ff}$, a simple
3-dimensional star formation model can be built with the form:
\begin{equation}
    \dot\rho_* = \frac{c_*\rho_{gas}}{t_{ff}} = c_*\sqrt{\frac{\rho^3_{gas}}{4\pi G}}
\end{equation}
\citet{Katz1992} used a $c_*$ parameter of $0.1$, which is within the range of
what is typically used in simulations today ($1-100\%$).  On top of the model
for star formation, feedback was included without the use of any significant
subgrid model:  $10^{51}\;\rm{erg/SN}$ worth of energy was injected back into
the ISM by smoothing it over a number of resolution elements.  One of the key
findings of \citet{Katz1992} was that the vast majority of SN energy was
radiated away, making them very ineffective at regulating star formation within
the galaxy.  In the end, \citet{Katz1992} found that the peak star formation in
the simulated galaxy was nearly 10 times greater than in observed galaxies.
\citet{Thacker2000} later showed that the smoothing of energy used by
\citet{Katz1992} resulted in a large overestimation of the cooling rates in
feedback-heated gas. Since then, much work has gone into developing models for
feedback that address the issue of overcooling that was present in these
earlier simulations.

@@

\subsection{Superbubbles: The Natural Scale for Feedback}
The formation of massive stars takes place mostly in clusters of $10^4$ or more
stars.  This means that the SN of these stars can be treated not as individual
dumps of $10^{51}\;\rm{erg/SN}$, but instead as a luminosity of
$10^{34}\;\rm{\frac{erg}{s \cdot M_\odot}}$.  This allows us to use stellar wind
models like those of \citet{Weaver1977} to study the evolution of superbubbles driven by
hot stellar feedback processes, like SN or stellar winds.  Each individual SN's
ejecta thermalizes in a small region within the center of the bubble, and the
resulting hot gas drives a shock that sweeps up the surrounding ISM.  In a short
time, the swept-up ISM can radiatively cool efficiently resulting in a thin,
cold shell surrounding a hot, diffuse bubble.

The structure and evolution of superbubbles have been the subject of a great
deal of analytic study.  \citet{Castor1975} found that the time evolution of
superbubbles has a similarity solution, the \citet{Sedov1959} adiabatic blast
wave solution: 
\begin{equation}
    R(t)=27n_0^{-1/5}L_{36}^{1/5}t_6^{3/5}\;\rm{pc} 
\end{equation}
\begin{equation}
    V(t)=16n_0^{-1/5}L_{36}^{1/5}t_6^{-2/5}\;\rm{km/s}
\end{equation}
$n_0$ is the ambient density in $\rm{cm^{-3}}$, $L_{36}$ is the luminosity of
the star cluster in $10^{36}\;\rm{erg/s}$, and $t_6$ is the time in
$10^6\;\rm{yr}$.  The evolution of hot mass in the bubble, and the effects of
cooling in the swept-up shell, were examined in \citet{Weaver1977}.  A major
result of this study was the realization that thermal conduction between the hot
bubble and cold shell will result in an evaporative mass flux, pulling material
into the hot bubble.  This evaporated mass will quickly dominate over the small
amount of ejecta from stellar winds (and later SN).  The \citet{Weaver1977}
results were extended to clustered SN by \citet{MacLow1988}.  They calculated
that the typical cooling time for the hot bubble was $\sim15\;\rm{Myr}$, with a
weak dependence on the total luminosity and ambient density.  Before significant
cooling can occur in the hot interior, $\sim35\%$ of the total feedback
luminosity is lost to cooling of the swept-up shell, which cools in
$\sim10^4\;\rm{yr}$.  This means that, rather than the $1-10\%$ of energy 
preserved from a single supernova blast \citep{Chevalier1974}, $\sim65\%$ of the
initial energy budget is retained in the kinetic energy of the cold shell and
thermal energy of the hot interior.  \citet{MacLow1988} also demonstrated that,
as superbubbles do not have surface tension, they are not depressurized by cold
clouds, and instead simply deform around them while continuing to expand.
\citet{Silich1996} studied the effects of an ISM permeated with cold clumps in
greater detail, finding that the general behaviour discussed in
\citet{MacLow1988} held even in the case of an inhomogeneous ISM.  Even with
$50\%$ of the ISM mass trapped in cold clouds (which penetrate the cold shell of
the superbubble), the total hot mass within the superbubble will only vary by
$\sim20\%$.
